Manuscript

### 2019-10-29 manuscript.Rmd
update manu w/ figs

---
output              :
  word_document     :
    reference_docx  : ./word_reference_template.docx
    toc             : true
    toc_depth       : 2
    fig_caption     : true
    df_print        : kable
    # citation_package: natbib
title               : "Manuscript Draft"
author              :
- Sally E. Claridge
- Daniel M. Charytonowicz
- Adam A. Margolin
date                : "`r format(Sys.time(), '%d %B %Y')`"
bibliography        : manuscript.bib
fontsize            : 12pt
geometry            : margin=1in
linkcolor           : blue
urlcolor            : blue
citecolor           : blue
floatsintext        : yes
figurelist          : no
tablelist           : no
footnotelist        : no
linenumbers         : yes
abstract            : Multiple high-throughput functional screens in cancer cell lines have generated large amounts of information on drug efficacy in a variety of genomic contexts and cancer lineages. Reported inconsistencies between datasets have led us to investigate whether these large-scale screens reproduce established clinical drug-gene associations and if genomic features particular to specific genes improve said reproducibility. We evaluated three large-scale, small-molecule drug screens within the context of clinical interpretations derived from a new cancer variant annotation resource published by the Variant Interpretation for Cancer Consortium (VICC). We identified low levels of concordance between the three drug screen datasets and clinical drug-gene associations using mutation status, gene expression, and copy number as genomic indicators. Less than half of the clinical drug-gene cancer associations from the VICC resource were identified in these three drug screen datasets, suggesting a barrier to translating findings from these large-scale screens into the clinic.
---

@@

@@

> Mutation status

> Fusions

> Copy number

> Gene expression

However, taking these results as a whole, we observe little concordance between G2P associations and the correlation of mutation status and AUC for said G2P association, suggesting that mutation status alone is not a robust predictor of clinically actionable genetic targets. This led us to examine the correlations between AUC and gene expression and AUC and copy number, two other genomic features that are known to associate with cancer and drug response. We computed Spearman correlation coefficients ($r_s$) between AUC of each drug-gene association and gene expression (RPKM) and copy number (log$_2$ ratio) of the gene in the association...**(???) need to complete this thought/analysis**

@@

- What do these correlations with mutation status, gene expression, and copy number look like when you constrict to a specific lineage?
- When you restrict, do the correlations become more distinct?
- Are the lineage-specific results approximately the same as the pancancer results?

@@

@@

To preclude variation in genomic feature calls across the datasets, we used identical annotations from CCLE for all cell lines screened in the three datasets. The CCLE provided mutation annotations in 19,280 genes across 1,596 cell lines (dataset: depmap_18Q4_mutation_calls.csv, accessed 14 November 2018). CCLE called substitutions using MuTect [@cibulskis2013] and annotated variants using Oncotator [@ramos2015a] and indels using Indelocator (https://software.broadinstitute.org/cancer/cga/indelocator). We filtered mutation calls for point mutations, defined as single-nucleotide insertions, deletions, and substitutions, regardless of result, i.e. frameshift, missense, or nonsense mutation. For this analysis, cell lines harboring non-silent point mutations were considered “mutant” for the gene in question. All other mutations and genes without annotation in the Mutation Annotation Format (MAF) file were considered “wildtype.” “Mutant” and “wildtype” calls were binarily encoded per-gene/per-cell line, regardless of quantity of harbored mutations per gene in a given cell line.

>>>>

---
output              :
  word_document     :
    reference_docx  : ./word_reference_template.docx
    toc             : true
    toc_depth       : 2
    fig_caption     : true
    df_print        : kable
    # citation_package: natbib
header-includes:
- \usepackage{caption}
title               : "Manuscript Draft"
author              :
- Sally E. Claridge
- Daniel M. Charytonowicz
- Adam A. Margolin
date                : "`r format(Sys.time(), '%d %B %Y')`"
bibliography        : manuscript.bib
fontsize            : 12pt
geometry            : margin=1in
linkcolor           : blue
urlcolor            : blue
citecolor           : blue
floatsintext        : yes
figurelist          : no
tablelist           : no
footnotelist        : no
linenumbers         : yes
abstract            : Multiple high-throughput functional screens in cancer cell lines have generated large amounts of information on drug efficacy in a variety of genomic contexts and cancer lineages. Reported inconsistencies between datasets have led us to investigate whether these large-scale screens reproduce established clinical drug-gene associations and if genomic features particular to specific genes improve said reproducibility. We evaluated three large-scale, small-molecule drug screens within the context of clinical interpretations derived from a new cancer variant annotation resource published by the Variant Interpretation for Cancer Consortium (VICC). We identified low levels of concordance between the three drug screen datasets and clinical drug-gene associations using mutation status, gene expression, and copy number as genomic indicators. Less than half of the clinical drug-gene cancer associations from the VICC resource were identified in these three drug screen datasets, suggesting a barrier to translating findings from these large-scale screens into the clinic.
---

@@

## Curating clinical biomarkers

@@

![Overlap of cancer cell lines in our four datasets.](./plots/manuscript/all_venn_ccls.png)

### Mutation status

![Mutation status.](./plots/manuscript/all_mut_heatmap.png)

![Protein changes.](./plots/manuscript/all_prot_heatmap.png)

Taking these results as a whole, we observe little concordance between G2P associations and the correlation of mutation status and AUC for said G2P association, suggesting that mutation status alone is not a robust predicto r of clinically actionable genetic targets. This led us to examine the correlations between AUC and gene expression and AUC and copy number, two other genomic features that are known to associate with cancer and drug response. We computed Spearman correlation coefficients ($r_s$) between AUC of each drug-gene association and gene expression (RPKM) and copy number (log$_2$ ratio) of the gene in the association...**need to complete this thought/analysis**

### Fusions

![Fusion status.](./plots/manuscript/all_fus_heatmap.png)

### Copy number

![Correlation between putative copy number and AUC.](./plots/manuscript/drug_cn_heatmap.png)

### Gene expression

![Correlation between gene expression and AUC.](./plots/manuscript/drug_ge_heatmap.png)

@@

- What do these correlations with mutation status, gene expression, and copy number look like when you constrict to a specific lineage?
- When you restrict, do the correlations become more distinct or are the lineage-specific results approximately the same as the pancancer results?

@@

**Need to update all methods for data processing, these method are current up to ~May 2019**

@@

To preclude variation in genomic feature calls across the datasets, we used identical annotations from CCLE for all cell lines screened in the three datasets. The CCLE provided mutation annotations in 19,280 genes across 1,596 cell lines (dataset: depmap_18Q4_mutation_calls.csv, accessed 14 November 2018). CCLE called substitutions using MuTect [@cibulskis2013] and annotated variants using Oncotator [@ramos2015a] and indels using Indelocator (https://software.broadinstitute.org/cancer/cga/indelocator). We filtered mutation calls for point mutations, defined as single-nucleotide insertions, deletions, and substitutions, regardless of result, i.e. frameshift, missense, or nonsense mutation. For this analysis, cell lines harboring non-silent point mutations were considered “mutant” for the gene in question. All other mutations and genes without annotation in the Mutation Annotation Format (MAF) file were considered "wildtype." "Mutant" and "wildtype" calls were binarily encoded per-gene/per-cell line, regardless of quantity of harbored mutations per gene in a given cell line.

### Protein changes

### Fusion status
